Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2014-02-02 background.tex
2013.07.28 (Vassilis) Further investigation conducted to find research work that references the papers/algorithms we survey and cite. Text enriched.

%
% TODO: SOME DISCUSSION ON P2P ARCHITECTURES
%
%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
%showed that the ad-hoc network topology of unstructured overlay networks that
%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
%Law describes the node degree while the Small World describes characteristics
%of path length and clustering coefficient. The clustering coefficient for a
%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
%the existing connections between $\upsilon$'s neighboring nodes to $\gamma
%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighboring
%nodes of $\upsilon$. High cluster coefficient means that neighboring nodes of
%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
%systems can have high maintenance traffic in delivering messages, updating the
%mapping, discovering failures and replicating lost data or pointers, making
%them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
%there is a tight control between the data placement and the topology of the
%network. For this reason it is hard to efficiently support partially matched
%queries while Gnutella and other similar systems effortlessly support keyword
%searches and other complex queries since the mechanism is realized locally, on
%a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely
%for a flooding-based search to return results. DHTs, on the other hand, fit
%better in the systems which require ability to reliably locate content, even in
%the extreme case that only a single-copy exists in the network.
%\end{itemize}

@@

The topology mismatch problem is exacerbated by the extremely large amounts of
traffic generated by highly popular \p\ applications~\cite{byl-book}.
Addressing this problem remains a fundamental challenge 
in contributing to the overall efficiency of the Internet. 
Unfortunately, topology mismatch is known to be 
an \emph{NP-Hard} problem \cite{C2000,NPBOOK}. 
There is also additional complexity due to the fact that 
end-to-end latencies demonstrate \emph{triangle inequality
violations (TIVs)} which are a consequence of the Internet's structure and
routing policies; as such, they will remain a property of the Internet 
for the foreseeable future \cite{zheng_irprtt_2005}. 
\emph{TIV}s affect both network coordinate
\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
systems and makes the construction of \emph{IP}-topology-aware 
overlays difficult.

>>>>

%
% TODO: SOME DISCUSSION ON P2P ARCHITECTURES
%
%\cite{matei_mapgnutella_2002, LCCLS2002, merugu_str2unstr_2003}
%showed that the ad-hoc network topology of unstructured overlay networks that
%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
%Law describes the node degree while the Small World describes characteristics
%of path length and clustering coefficient. The clustering coefficient for a
%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
%the existing connections between $\upsilon$'s neighboring nodes to $\gamma
%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighboring
%nodes of $\upsilon$. High cluster coefficient means that neighboring nodes of
%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
%systems can have high maintenance traffic in delivering messages, updating the
%mapping, discovering failures and replicating lost data or pointers, making
%them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
%there is a tight control between the data placement and the topology of the
%network. For this reason it is hard to efficiently support partially matched
%queries while Gnutella and other similar systems effortlessly support keyword
%searches and other complex queries since the mechanism is realized locally, on
%a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely
%for a flooding-based search to return results. DHTs, on the other hand, fit
%better in the systems which require ability to reliably locate content, even in
%the extreme case that only a single-copy exists in the network.
%\end{itemize}

@@

The topology mismatch problem is exacerbated by the extremely large amounts of
traffic generated by highly popular \p\ applications~\cite{byl-book}.
Addressing this problem remains a fundamental challenge 
in contributing to the overall efficiency of the Internet. 
Unfortunately, topology mismatch is known to be 
an \emph{NP-Hard} problem \cite{C2000,NPBOOK}. 
There is also additional complexity due to the fact that 
end-to-end latencies demonstrate \emph{triangle inequality
violations (TIVs)} which are a consequence of the Internet's structure and
routing policies; as such, they will remain a property of the Internet 
for the foreseeable future \cite{zheng_irprtt_2005}. 
\emph{TIV}s affect both network coordinate
\cite{cox_vivaldi_2004,WSS2005} and positioning \cite{ng_gnp_2001}
systems and makes the construction of \emph{IP}-topology-aware 
overlays difficult.

### 2014-02-02 structured.tex
2013.07.28 (Vassilis) Further investigation conducted to find research work that references the papers/algorithms we survey and cite. Text enriched.

\emph{Global Soft-State} \cite{XTZ2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network. %Figure \ref{fig:global_softstate} presents the landmark
%space and the mapping of the position of the nodes.
This global view of the
state, is made available to all nodes to help them find the best way to
route their messages. \emph{Global Soft-State} operates in two phases:
generation and use of the proximity information. For generating the proximity
information, a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step to select a number of potential nearest neighbor
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm follows a different path from that of the
classic gossiping approaches for the construction and maintenance of the
overlay. It is based on landmark clustering for generation and strategic
placement of proximity information on the overlay. This information is called a
``map'' and concerns a region of the overlay (part of the Cartesian space in
structures like eCAN\cite{xu_ecan_2002} or a set of nodes sharing a particular
prefix in overlays such as Pastry\cite{antony_pastry_2001}). Thus, each node not
only can access such information from these landmarks but can also subscribe to
relevant \emph{soft states} and get notified on dynamic network changes. A
node may appear in a maximum of $log\left( N \right)$ such
maps.
%%
Although this approach greatly reduces the routing latency to far nodes, it
comes expensive. It is possible to end up with a very large number of regions
each of which has to maintain a map. The fact that a peer may appear in multiple
such maps we can conclude that additional traffic for probing and notifications
renders the approach does not scalable. Moreover, to refine measurements
additional inner-bin measurement s are required\cite{WZS2004}. Also improvement
is done within a region because of little knowledge about neighboring zones.

@@

One major limitation of \emph{Mithos} is that the protocol cannot efficiently handle
dynamic arrivals and departures from the network. Moreover, parts of this spring
relaxation technique requires a centralized
implementation~\cite{cox_vivaldi_2004}.

%% VM I have read this somewhere but I didn't keep the reference ... :(
%%
% In order to avoid local minima during neighbor detection, extensive probing
% must be undertaken. In simulation, unfortunately, only very small-sized
% overlay topologies (of 200 to 1000 nodes) have been used and thus no safe
% conclusions can be made as for the behavior of an extensively large,
% real-world \p\ deployment of the scheme.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% The efficiency of choosing close by peers is very good and reported almost
% 65\% better than that of GNP (and thus generally of static landmarking as a
% whole)
high &
% trace-route is a very heavy tool and thus incurs excessive burden to the
% network resources.
high &
% Trace-route is categorized as too heavy-weighted and intrusive for use in a
% larger scale system \cite{RHMKS2002}. Additionally, disabling
% ICMP is a common administrative policy for edge sites to enforce security,
% while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not
% directly available to the application layer.
low
\end{tabular}
\end{center}

@@

\emph{PChord}’s advantages include the fact that the main routing optimizations are of
less overlay hops, following proximity links and crossing the same network
partition only once along a routing path while in the same time keeping Chord's
light-weight maintenance cost. On the other hand, it exhibits slow convergence
speed and inefficiency in a high churn environment\cite{DK2006}.

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Small decrease in needed overlay hops per object publishing but rigid in
% high churn environments.
medium &
% The are heartbeat mechanism for both finger table and proximity list adds a
% burden to the network by default. But the overall overhead is decreased as
% the proximity list becomes longer.
medium &
% Improvement of routing is achieved with more object pointers helping to
% shorten the routing path of object querying. When the distance of query source
% to target is far away both Chord and PChord achieve the average hop number to
% not increase.
high
\end{tabular}
\end{center}

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Can reduce inter-domain average path length but even though nodes in the same
% domain would have close identifiers, nodes in two close domains may have very
% different ones reducing the overall efficiency of the approach in general when
% considering inter-domain average path length. No active adaptation mechanism
% for high churn environments.
low &
% The construction of the identifier space is done at node join and in an
% independent fashion involving the application of two hash functions meaning
% resulting in low overhead stress to the network and its nodes.
low &
% Is designed on the basis of IPv6 protocol which is not widely operational yet.
medium
\end{tabular}
\end{center}

@@

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% The efficiency of proximity detection through the prefix matching technique
% is caught in a vicious cycle with the storage space needed for DHT object
% publishing (how many bits will be used for either)
low &
% The overhead for detecting proximity is only comparison of prefixes
low &
%
medium
\end{tabular}
\end{center}

@@

\begin{center}
{\footnotesize
\begin{tabular}{rccc}
\multicolumn{1}{r}{} &
\multicolumn{1}{c}{\emph{Efficiency}} &
\multicolumn{1}{c}{\emph{Overhead}} &
\multicolumn{1}{c}{\emph{Scalability}}
\\
\cline{2-4}
\emph{RLM} &
%
medium &
%
medium &
% Dynamic landmarking approach helps the scalability of the algorithm. The
% authors claim that the temporary nature of RLM's landmark nodes does not
% prevent them to maintain a good ratio in the presence of high landmark
% degression rates.
high \\
\emph{CNPA} &
%
medium &
% % The paper suggests that this approach generates significantly lower overhead
% than the vanilla Pastry algorithm
low &
% The queries are directed through fast nodes so it is important such nodes
% be part of the network
medium \\
\end{tabular}
}
\end{center}

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% SAT-Match seems to outperform landmark binning in the sense of providing a
% higher grained differentiation of proximity, adding to the fact that
% SAT-match is fully distributed. Also to further enhance the efficiency it can
% work cooperatively with landmark binning itself.
high &
% distance probing and iterative computations add to the approach's overhead.
% Also the fact that the selective jumps are not always performed the added
% overhead to efficiency gained is not guaranteed in any configuration and
% circumstance.
medium &
% The self-adaptation nature of the algorithm and the selective jump mechanisms
% scale much better than, for example, Mithos. The test are on a relatively
% small amount of peers, thus, no safe conclusion can be made with respect to the
% algorithm's scalling performance.
medium
\end{tabular}
\end{center}

@@

Even though managing the overlay structure based on the \emph{GL} of the
nodes improves the query efficiency of the system, on the other hand, it tends
to create hot-spots, and the needed failure resilience is undermined by the fact
that nearby nodes are more likely to suffer collective failures.

>>>>

\emph{Global Soft-State} \cite{XTZ2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network. %Figure \ref{fig:global_softstate} presents the landmark
%space and the mapping of the position of the nodes.
This global view of the
state, is made available to all nodes to help them find the best way to
route their messages. \emph{Global Soft-State} operates in two phases:
generation and use of the proximity information. For generating the proximity
information, a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step to select a number of potential nearest neighbor
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm follows a different path from that of the
classic gossiping approaches for the construction and maintenance of the
overlay. It is based on landmark clustering for generation and strategic
placement of proximity information on the overlay. This information is called a
``map'' and concerns a region of the overlay (part of the Cartesian space in
structures like eCAN\cite{xu_ecan_2002} or a set of nodes sharing a particular
prefix in overlays such as Pastry\cite{antony_pastry_2001}). Thus, each node not
only can access such information from these landmarks but can also subscribe to
relevant \emph{soft states} and get notified on dynamic network changes. A
node may appear in a maximum of $log\left( N \right)$ such
maps.
%%
Although this approach greatly reduces the routing latency to far nodes, it
comes expensive. It is possible to end up with a very large number of regions
each of which has to maintain a map. The fact that a peer may appear in multiple
such maps we can conclude that additional traffic for probing and notifications
renders the approach not scalable \cite{RGJZ2004}. Moreover, to refine measurements
additional inner-bin measurement s are required\cite{WZS2004}. Also improvement
is done within a region because of little knowledge about neighboring zones.

@@

One major limitation of \emph{Mithos} is that the protocol cannot effectively
handle dynamic arrivals and departures from the network. In \cite{CCRK2004} it
is claimed to need extensive probing to determine distances and at the same time
its results are evaluated as less accurate than, for example,
GNP \cite{ng_gnp_2001}. Maybe this has also to do with its iterative neighbor
selection process which is reported to be prone to prematurely terminate at
local minima \cite{WSS2005}. Another point that has negative effect to the
scalability of the algorithm is that parts of its spring relaxation technique
require a centralized implementation~\cite{cox_vivaldi_2004}.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% The efficiency of choosing close by peers is very good and reported almost
% 65\% better than that of GNP (and thus generally of static landmarking as a
% whole)
high &
% trace-route is a very heavy tool and thus incurs excessive burden to the
% network resources.
high &
% Trace-route is categorized as too heavy-weighted and intrusive for use in a
% larger scale system \cite{RKHS2002}. Additionally, disabling
% ICMP is a common administrative policy for edge sites to enforce security,
% while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not
% directly available to the application layer.
low
\end{tabular}
\end{center}

@@

\emph{PChord}’s advantages include the fact that the main routing optimizations are of
less overlay hops, following proximity links and crossing the same network
partition only once along a routing path while in the same time keeping Chord's
light-weight maintenance cost. On the other hand, it exhibits slow convergence
speed and inefficiency in a high churn environment \cite{DK2006}.

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Small decrease in needed overlay hops per object publishing but rigid in
% high churn environments.
medium &
% The are heartbeat mechanism for both finger table and proximity list adds a
% burden to the network by default. But the overall overhead is decreased as
% the proximity list becomes longer.
medium &
% Improvement of routing is achieved with more object pointers helping to
% shorten the routing path of object querying. When the distance of query source
% to target is far away both Chord and PChord achieve the average hop number to
% not increase.
medium
\end{tabular}
\end{center}

@@

This approach is able to reduce the average path length of the system to be
logarithmically proportional to the number of available domains instead of the
number of participating peers. However due to the large number of Internet
domains and the potentially small number of system peers within each domain,
this approach doesn't seem scalable and efficient \cite{DK2006}.

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Can reduce inter-domain average path length but even though nodes in the same
% domain would have close identifiers, nodes in two close domains may have very
% different ones reducing the overall efficiency of the approach in general when
% considering inter-domain average path length. No active adaptation mechanism
% for high churn environments.
low &
% The construction of the identifier space is done at node join and in an
% independent fashion involving the application of two hash functions meaning
% resulting in low overhead stress to the network and its nodes.
low &
% Is designed on the basis of IPv6 protocol which is not widely operational yet.
low
\end{tabular}
\end{center}

@@

The algorithm's overhead is very low because detection of proximity is only
a comparison of prefixes. On the other hand, the efficiency of the proximity
is caught into a vicious cycle with the storage space needed for DHT object
publishing (how many bits will be used for either). Moreover, its static
nature (no online measurements used) makes it difficult to handle
\cite{EWF2011}.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% The efficiency of proximity detection through the prefix matching technique
% is caught in a vicious cycle with the storage space needed for DHT object
% publishing (how many bits will be used for either).
low &
% The overhead for detecting proximity is only comparison of prefixes
low &
%
medium
\end{tabular}
\end{center}

@@

\begin{center}
{\footnotesize
\begin{tabular}{rccc}
\multicolumn{1}{r}{} &
\multicolumn{1}{c}{\emph{Efficiency}} &
\multicolumn{1}{c}{\emph{Overhead}} &
\multicolumn{1}{c}{\emph{Scalability}}
\\
\cline{2-4}
\emph{RLM} &
%%
%% TAKEN FROM
%% Efficient service discovery mechanism for wireless sensor networks
%%
%%
% common prefix of the overlay ID with the closest landmark. The
% RLM utilizes the overlay lookup capabilities by reducing traffic
% overhead within an overlay cluster. However, it generates much
% control traffic for periodically measuring distances to all landmarks.
% Moreover, the hop stretch overhead still increases when
% performing overlay routing among overlay clusters.
medium &
%
medium &
% Dynamic landmarking approach helps the scalability of the algorithm. The
% authors claim that the temporary nature of RLM's landmark nodes does not
% prevent them to maintain a good ratio in the presence of high landmark
% degression rates.
high \\
\emph{CNPA} &
%
medium &
% % The paper suggests that this approach generates significantly lower overhead
% than the vanilla Pastry algorithm
low &
% The queries are directed through fast nodes so it is important such nodes
% be part of the network
medium \\
\end{tabular}
}
\end{center}

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% SAT-Match seems to outperform landmark binning in the sense of providing a
% higher grained differentiation of proximity, adding to the fact that
% SAT-match is fully distributed. Also to further enhance the efficiency it can
% work cooperatively with landmark binning itself.
% Security is compromised with arbitrary jumps!
high &
% distance probing and iterative computations add to the approach's overhead.
% Also the fact that the selective jumps are not always performed the added
% overhead to efficiency gained is not guaranteed in any configuration and
% circumstance.
medium &
% The self-adaptation nature of the algorithm and the selective jump mechanisms
% scale much better than, for example, Mithos. The test are on a relatively
% small amount of peers, thus, no safe conclusion can be made with respect to the
% algorithm's scalling performance.
medium
\end{tabular}
\end{center}

@@

Even though managing the overlay structure based on the \emph{GL} of the
nodes improves the query efficiency of the system, on the other hand, it tends
to undermine the uniform distribution of peer identifiers which consequently
tend to create hot-spots, failing resilience is undermined by the fact
that nearby nodes are more likely to suffer collective failures \cite{HY2007}.

% must
% cope with the inherent trade os associated with selecting
% identiers based on location: randomly selected identiers
% are essential for the scalability and security of the system
% and help to balance load while location-aware identiers pre-
% vent messages from taking extensive scenic routes (on the
% physical level).
